---
title: "R Notebook"
output: html_notebook
---



#Problem 2
```{r}
set.seed(12345)
y <- seq(from=1, to=100, length.out = 1e5) + rnorm(1e5)
```




```{r}
system.time({
  ss.2 <- 0
  for(i in 1:length(y)){
    ss.2 <- ss + (y[i] - mean(y))^2
  }
  
})

system.time({
ss.1 <- sum((y - mean(y)^2))  
})
```


##Problem 3 
```{r}
set.seed(1256)
theta <- as.matrix(c(1,2),nrow=2)
X <- cbind(1,rep(1:10,10))
h <- X%*%theta+rnorm(100,0,0.2)
m <- nrow(h)
theta.0 <- vector()
theta.1 <- vector()
theta.0[1] <- 0
theta.1[1] <- 0
theta.0[2] <- 3
theta.1[2] <- 5
alpha <- .01
h.o <- vector()
i <-2
tolerance <- .00000000000001
while (abs(theta.0[i]-theta.0[i-1]) > tolerance & abs(theta.1[i]-theta.1[i-1]) > tolerance) {
    if(i > 100000){
      break
    } else{
      h.o <- theta.0[i] + theta.1[i]*X[,2]   
    theta.0[i+1] <- theta.0[i] - alpha*(1/m)*sum(h.o - h)
    theta.1[i+1] <- theta.1[i] - alpha*(1/m)*sum((h.o - h)*X[,2])
    }
  i <- i+1
  }
  
theta.0[length(theta.0)]
theta.1[length(theta.1)]
```

```{r}
lm(h~0+X)
```


The coefficents obtained from the while loop are very similar to the coefficents obtained from the lm function, the tolerance I used is 1e-14 with a step size of .01

#Problem 4

John Cook suggested that we should not solve for the inverse but instead solve $(X^TX)\beta = X^Ty$. This would be done in R using the solve function.


#Problem 5
```{r}
rm(list=ls())
set.seed(12456) 
G <- matrix(sample(c(0,0.5,1),size=16000,replace=T),ncol=10)
R <- cor(G) # R: 10 * 10 correlation matrix of G
C <- kronecker(R, diag(1600)) # C is a 16000 * 16000 block diagonal matrix
id <- sample(1:16000,size=932,replace=F)
q <- sample(c(0,0.5,1),size=15068,replace=T) # vector of length 15068
A <- C[id, -id] # matrix of dimension 932 * 15068
B <- C[-id, -id] # matrix of dimension 15068 * 15068
p <- runif(932,0,1)
r <- runif(15068,0,1)
C<-NULL #save some memory space

object.size(A)
object.size(B)


system.time({
  y = p + A %*% (solve(B) %*%(q - r))
})


system.time({
  qr <- q-r
  B.1 <- solve(B)
  B.1qr <- (B.1%*%qr)
  rm(B.1,qr)
  ABqr <- A%*%B.1qr
  rm(A)
  y = p + ABqr
})

system.time({
  qr <- q-r
  B.1 <- solve(B)
  AB <- crossprod(A,B)
  ABQR <- crossprod(AB,qr)
  y = p + ABQR
})



system.time({
  t(solve(B %*% solve((q-r)%*%t(q-r))%*%(q-r),A)+t(p))
})
```
##Problem 5 Part A
The A matrix is 112347208 Mb and the B matrix 1816357192 bytes.

##Problem 5 Part B
I tried solving the equation to get:

$B[(q-r)(q-r)^T]^{-1}(q-r)(y^T-p^T) = A$

and then using the solve function one could obtain:

$y = solve(B[(q-r)(q-r)^T]^{-1}(q-r),A)+p^T = A$

I thought in theory this might be easier to solve but I had no luck.



